---
title: "Analytics Capstone"
author: "Jai Prakash"
date: "10/30/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("e1071")
library(e1071) 
library(DataExplorer)
library(dplyr) 
library(tidyverse)
library(lubridate)
library(plotly)
library(forecast)
library(xts)

#Function for Mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
options(scipen = 999)
#options(scipen = 0)
```

Loading the Data from data.csv

```{r , include=FALSE}
setwd("C:/Docs/Learning/iMBA/8 Analytics/Capstone")
trans <- as.data.frame(read.csv("data.csv"))
attach(trans)
```

```{r}
#Viewing data Structure
str(trans)
introduce(trans)
head(trans)

#backup
transbkp <- trans
#trans <- transbkp



```
We have 8 variables in the data set, which includes 5 which can be factors, 1 date and 2 values.
We need to convert InvoiceDate as a Date and CustomerID as a factor for further analysis

```{r}
#Transform variables
trans <- trans %>% mutate(InvoiceDate=as.Date(InvoiceDate, '%m/%d/%Y %H:%M'),CustomerID=as.factor(CustomerID),Country=as.factor(Country),ItemSalesPrice=Quantity*UnitPrice)
str(trans)
introduce(trans)
```

#Univariate Analysis
## Continous Variables - Quantity and UnitPrice
```{r}

#Quantity
## Central Tendencies
summary(trans$Quantity)
getmode(trans$Quantity)

## Dispersion Measure
range(trans$Quantity)
quantile(trans$Quantity)
IQR(trans$Quantity)
var(trans$Quantity)
sd(trans$Quantity)
skewness(trans$Quantity)
kurtosis(trans$Quantity)


#UnitPrice
## Central Tendencies
summary(trans$UnitPrice)
getmode(trans$UnitPrice)

## Dispersion Measure
paste("range",range(trans$UnitPrice))
paste("quantile",quantile(trans$UnitPrice))
paste("IQR",IQR(trans$UnitPrice))
paste("Variance",var(trans$UnitPrice))
paste("SD",sd(trans$UnitPrice))
paste("skewness",skewness(trans$UnitPrice))
paste("kurtosis",kurtosis(trans$UnitPrice))
#hist(trans$UnitPrice,breaks = 10)
```

##Visualization - Univariate Analysis
```{r}
plot_boxplot(trans, by="Country",geom_boxplot_args = list("outlier.color" = "red"))
hist(trans$Quantity[which(abs(trans$Quantity)<100)])
hist(trans$UnitPrice[which(abs(trans$UnitPrice)<100)])
plot_histogram(trans[which(abs(trans$Quantity)<100),],geom_histogram_args = list(bins = 30L))
plot(trans$Quantity,trans$UnitPrice)
```
#Univariate Analysis
## Categorical Variables - Country, InvoiceNo, StockCode, Description, CustomerID
```{r}

#Country
#trans$Country
table(trans$Country)
table(trans$Country)/nrow(trans)
plot_bar(trans$Country)
plot_bar(trans$Country[-which(trans$Country =='United Kingdom')])
sort(table(trans$StockCode),decreasing=TRUE)[1:10]
sort(round(prop.table(table(trans$StockCode))*100,digits=4),decreasing=TRUE)[1:10]


#Top10 values
##Country
sort(table(trans$Country),decreasing=TRUE)[1:10]
sort(round(prop.table(table(trans$Country))*100,digits=4),decreasing=TRUE)[1:10]

##StockCode
length(unique(StockCode))
sort(table(trans$StockCode),decreasing=TRUE)[1:10]
sort(round(prop.table(table(trans$StockCode))*100,digits=4),decreasing=TRUE)[1:10]
###Discounts
count(trans[Description=="Discount",])


##Description
sort(table(trans$Description),decreasing=TRUE)[1:10]
sort(round(prop.table(table(trans$Description))*100,digits=4),decreasing=TRUE)[1:10]

##CustomerID
length(unique(CustomerID))
sort(table(trans$CustomerID),decreasing=TRUE)[1:10]
sort(round(prop.table(table(trans$CustomerID))*100,digits=4),decreasing=TRUE)[1:10]

##InvoiceDate
str(trans)
class(InvoiceDate)
#yyyymm=make_date(year = year(InvoiceDate),month = month(InvoiceDate))
#sort(table(yyyymm),decreasing=TRUE)[1:10]
#sort(round(prop.table(table(yyyymm))*100,digits=4),decreasing=TRUE)[1:10]
#unique(substr(InvoiceDate,5,10))
#str(trans)

##InvoiceID
length(unique(InvoiceNo))

Cancelled<- ifelse(substr(InvoiceNo,1,1)=="C","Y","N")
trans<-cbind(trans,Cancelled)
prop.table(table(Cancelled))

#trans <- trans[,1:9]
```
We have about 24% of records which are missing Customer ID. These can be removed as they will not help with the customer segmentation analysis

#Bi-Variate Analysis

```{r}
# Continuous-Continuous Variable
plot(trans$ItemSalesPrice,trans$Quantity)
cor(trans$ItemSalesPrice,trans$Quantity)
```
```{r}

# Continuous-Categorical
#Sales by Invoice
trans.Inv <- trans %>% 
  group_by(InvoiceNo) %>% 
  summarise(TotalPrice=sum(ItemSalesPrice),TotalQty=sum(Quantity))
t<- trans.Inv

plot(t$TotalPrice,t$TotalQty)
summary(t)
getmode(t$TotalQty)
getmode(t$TotalPrice)

#Sales by Customer
trans.Cust <- trans %>% 
  group_by(CustomerID) %>% 
  summarise(TotalPrice=sum(ItemSalesPrice),TotalQty=sum(Quantity))
t<- trans.Cust

plot(t$TotalPrice,t$TotalQty)
summary(t)
getmode(t$TotalQty)
getmode(t$TotalPrice)

#Sales by Product
trans.Prod <- trans %>% 
  group_by(StockCode) %>% 
  summarise(TotalPrice=sum(ItemSalesPrice),TotalQty=sum(Quantity))
t<- trans.Prod

plot(t$TotalPrice,t$TotalQty)
summary(t)
getmode(t$TotalQty)
getmode(t$TotalPrice)

#Sales by Country
trans.Country <- trans %>% 
  group_by(Country) %>% 
  summarise(TotalPrice=sum(ItemSalesPrice),TotalQty=sum(Quantity),CustCount=length(unique(CustomerID)))
t<- trans.Country

plot(t$TotalPrice,t$TotalQty)
plot(t$Country,t$CustCount,type="h")
summary(t)
getmode(t$TotalQty)
getmode(t$TotalPrice)

```
```


```{r}

#Missing Value
plot_missing(trans)

#Missing Description
count(trans[Description=="",])

#Negative Price
count(trans[UnitPrice<=0.01,])

#Negative Quantity but not Cancelled items
count(trans[Quantity<=0 & Cancelled=="N",])

#Invalid Descriptions
length(unique(trans$Description[str_length(Description)<=8]))

###Non-Numeric Stock Codes
sort(unique(trans$StockCode),decreasing=TRUE)[1:33]


#Box Plot for Outliers
plot_boxplot(trans, by="Country",geom_boxplot_args = list("outlier.color" = "red"))
```


```{r}
#Data Cleanup
#Remove missing customer repords
trans<- trans[-which(is.na(trans$CustomerID)),]
#summary(trans$CustomerID)

#Remove records which are outliers
trans[which(abs(trans$Quantity)>50000),]
trans<- trans[-which(abs(trans$Quantity)>50000),]

head(trans[which(trans$Quantity<0),])
head(trans[which(trans$UnitPrice<0.1),])

trans[which(abs(trans$UnitPrice)>5000),]
trans<- trans[-which(abs(trans$UnitPrice)>5000),]

#trans[which(abs(trans$Quantity)>20000),]
#trans[which(abs(trans$UnitPrice)>5000),]
#sum(trans$Quantity[which(abs(trans$UnitPrice)==0)])
#sum(trans$Quantity)
```


